---
title: "Covid-19's effect on crime in Toronto neighbourhoods"
subtitle: ""
author: 
  "Varun Vijay"
thanks: "Code and data are available at: https://github.com/Varun1005473462/final_folder-main.git."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---


```{r setup, include=FALSE}
library(tidyverse)
library(modelsummary)
library(readr)
library(here)
library(bookdown) # Referencing and report making
library(dplyr) # Data manipulation
library(kableExtra) #Table displaying
library(ggthemes) # More plot themes
library(reshape2) # Dataframe merging
library(janitor) # Cleaning data
library(ggplot2) # To make graphs
library(modelsummary)
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

After the first reported outbreak of covid-19 in December 2019, the entire world changed. Covid-19 was quickly declared to be a health emergency of global concern by the World health organization. Governments started imposing lockdowns, travel was banned,  public gatherings were restricted, buisnesses were closed down and everyone started isolating at home. Citizens of countries were essentially restricted to their homes and rarely left. Covid-19 had a huge impact on the global economy and has restricted people's actions and movements. The question this paper seeks to answer is how the covid pandemic has affected crime in various neighborhoods in Toronto. 

In this paper, we look at how the number of covid cases in a neighborhood can affect the various crime rates in the neighborhood such as assault rates and auto thefts. We construct a model using R based on the dataset in which we can see how the number of covid cases can affect the various crime statistics. We also use the dataset to construct various histograms, bar plots and summary statistics about the dataset to improve the quality of our analysis. The results obtained seem to imply that the covid pandemic has had a negative effect on crime in 2020.

The remainder of this paper is structured as follows: Section 2 will discuss the data for this paper, Section 3 will discuss the model that has been obtained, Section 4 will present the results obtained from the analysis of the dataset, and finally Section 5 will discuss the strengths and weaknesses of the paper.

# Data

```{r dataextract, echo=FALSE}
options(readr.show_col_types = FALSE)
covid_cases<-
  read_csv(here::here("inputs/data/covid_cases.csv"))
neighbourhood_crime_rates<-
  read_csv(here::here("inputs/data/neighbourhood_crime_rates.csv"))
covid_cases
neighbourhood_crime_rates
```
```{r}
covid_cases_neighbourhood<-data.frame(table(covid_cases['Neighbourhood Name']))
colnames(covid_cases_neighbourhood)[1]<-"Neighbourhood"
colnames(covid_cases_neighbourhood)[2]<-"Number of Covid cases"
covid_cases_neighbourhood 
```

```{r}
neighbourhood_crime_rates |>
  arrange(Neighbourhood)
covid_cases_neighbourhood |>
  arrange(Neighbourhood)
```


# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results




# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


